\vspace{-0.8in}Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms

### 2016-02-12 podc16/paper2.tex
spotted a mistake in graph processing. Left a few comments.

We now present our abstract leaderless generalized consensus algorithm.
A process $p$ receiving a command $c$ proceeds in three phases. First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $cs$ that the command $c$ may depend on. Second, the process $p$ proposes the mapping  $c\mapsto cs$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto cs'$ for $c$, in which case $cs'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable. 
Finally, when $c$ is executable, the process $p$ may execute $c$ by deciding the c-struct obtained by running the \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.

@@

\subsection{The Abstract Generalized Consensus Algorithm}
\label{abstract-algo}
We now propose describe our abstract generalized consensus algorithm. 
Local variables: $m_p$, a map from commands to dependency sets, initially the empty map; $\sigma_p$, a c-struct, initially $\bot$.
\begin{itemize}[noitemsep,nolistsep]
    \item Upon ${\opn{propose}\left( c \right)}^p$: call ${\opn{announce}\left( c \right)}^p$.
    \item Upon ${\opn{suggest}\left( c,d \right)}^p$: call ${\opn{propose}\left( c,d \right)}^p$.
    \item Upon ${\opn{decide}\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map $m_p$.
    \item Periodically, update $\sigma_p$ to $\opn{c-struct}\left( m_p \right)$, where $\opn{c-struct}\left( m_p \right)$ is defined below.
\end{itemize}

The graph processing algorithm is executed locally by a node. It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined below.
Its main property is Lemma 1; together with Invariants 1 and 2 (TODO, the invariants guaranteed by the dependency-set algorithm and the map-agreement algorithm), Lemma 1 implies that leaderless algorithms safely implement generalized consensus. 
We assume that processes initially agree, for each set of commands $cs$, on a total order $<_{cs}$ on $cs$.
In practice, if commands are attached a unique identifier among a totally ordered set, $<_{cs}$ can be the restriction of this total order on $cs$.

@@

\textbf{Lemma 1}
Assume that: $g_1$ and $g_2$ are two dependency graphs; 
for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$;
$g_1 \leq g_2$;
$l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
Then we have (a), $\bot\bullet l_1 \sqsubseteq \bot\bullet l_2$, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.

@@

Define the set of linearizations $\opn{Lins}\left( g \right)$ of a graph $g$ a the set of sequences whose length is the cardinality of $V\left( g \right)$, which contain each element of $V\left( g \right)$ exactly once, and ... \todo{or first define the partial order induced by a graph and then its compatible total orders.}

@@

\subsection{Correctness Proof}

@@

>>>>

We now present our abstract leaderless generalized consensus algorithm.
A process $p$ receiving a command $c$ proceeds in three phases. First, it calls a \textit{dependency-set} sub-algorithm (\Cref{dep-algo}) to determine a set of commands $cs$ that the command $c$ may depend on. Second, the process $p$ proposes the mapping  $c\mapsto cs$ to a \textit{map-agreement} sub-algorithm (\Cref{map-algo}); the map-agreement sub-algorithm may subsequently commit a mapping $c\mapsto cs'$ for $c$, in which case $cs'$ becomes the set of dependencies of $c$, and we say that $c$ is executable when its dependency set is empty or when all its dependencies are executable. 
Finally, when $c$ is executable, the process $p$ may execute $c$ by deciding the c-struct obtained by running the \emph{graph-processing} algorithm described in \Cref{abstract-algo}.
Any other process can run the three phases independently of $p$ and eventually execute the command despite a crash of $p$.
However, we show that a leaderless  generalized consensus algorithm is always safe and is live when its underlying dependency-set and map-agreement algorithms are safe and live.
The leaderless generalized consensus algorithm orchestrates the two dependency-set and map-agreement sub-algorithms as shown in \cref{fig:deps-set-algo3}.

@@

\subsection{Processing the Dependency Graph}
\label{abstract-algo}

\begin{comment}
Local variables: $m_p$, a map from commands to dependency sets, initially the empty map; $\sigma_p$, a c-struct, initially $\bot$.
\begin{itemize}[noitemsep,nolistsep]
    \item Upon ${\opn{propose}\left( c \right)}^p$: call ${\opn{announce}\left( c \right)}^p$.
    \item Upon ${\opn{suggest}\left( c,d \right)}^p$: call ${\opn{propose}\left( c,d \right)}^p$.
    \item Upon ${\opn{decide}\left( c,d \right)}^p$: insert the mapping $c \rightarrow d$ in the map $m_p$.
    \item Periodically, update $\sigma_p$ to $\opn{c-struct}\left( m_p \right)$, where $\opn{c-struct}\left( m_p \right)$ is defined below.
\end{itemize}

The graph processing algorithm is executed locally by a node (line 10 of \cref{fig:deps-set-algo3}). It must compute the value of $\opn{c-struct}\left( g_p \right)$, as defined below.
Its main property is Lemma 1, which with the properties of the sub-algorithms imply safety and liveness. 
We assume that processes initially agree, for each set of commands $cs$, on a total order $<_{cs}$ on $cs$.
In practice, if commands are attached a unique identifier among a totally ordered set, $<_{cs}$ can be the restriction of this total order on $cs$.

@@

\todo{Partial orders may not be what we need (see Lemma 2), instead:}
Define the intersection of two graphs as $V\left( g_1 \cap g_2 \right)=V\left( g_1 \right)\cap V\left( g_2 \right)$ and $E\left( g_1 \cap g_2 \right)= E\left( g_1 \right)\cap E\left( g_2 \right)$, and define their union similarly.
Two graphs $g_1$ and $g_2$ are \emph{compatible} iff $g_1 \cap g_2$ is a vertex-induced subgraph of $g_1 \cup g_2$.
\todo{Does this make any sense?}

\textbf{Lemma 1}
Assume that: $g_1$ and $g_2$ are two dependency graphs; 
for every pair of non-commutative commands $c_1,c_2 \in V\left(g_2\right)$, either $\left( c_1,c_2 \right)\in E\left( g_2 \right)$ or $\left( c_2,c_1 \right)\in E\left( g_2 \right)$;
$g_1 \leq g_2$; \todo{Instead of the previous one: $g_1$ and $g_2$ are compatible};
$l_1$ is a linearization of $g_1$ and $l_2$ is a linearization of $g_2$. 
Then we have (a), $\bot\bullet l_1 \sqsubseteq \bot\bullet l_2$, and (b), if $g_1 = g_2$ then $\bot\bullet l_1 = \bot\bullet l_2$.

@@

Define the set of linearizations $\opn{Lins}\left( g \right)$ of a graph $g$ as the set of sequences whose length is the cardinality of $V\left( g \right)$, which contain each element of $V\left( g \right)$ exactly once, and ... \todo{or first define the partial order induced by a graph and then its compatible total orders.}

@@

\subsection{Consistency and Liveness}
\textbf{Lemma 2}
If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1 \leq g_q^2$.
\todo{This is wrong\ldots}
What about: If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, and $t_1 \leq t_2$, then $g_p^1$ and $g_q^2$ are compatible.

@@

\appendix

\section{Generalized Consensus}
